### 2017-08-03 sheets/talk-6/sheet-6.tex
Merge branch 'uds-psl/master'

\begin{answer}
  The coequalizer follows the scheme like in the talk.
  We define the relation $R := \set{ (1_X (x), f(x) \;\big|\; x \in X} = \set{ (x, f(x) \;\big|\; x \in X} = f$ defined as a set.
  Now $\sim$ is the reflexive symmetric transitive closure of $R$.
  We define $i(x) := [x] := \set{ x' \in X | x \sim x'}$
  and the quotient $C := X / \sim \;=\; i(X)$.
  Now
  $ \xymatrix{
    X \ar@<.5ex>[r]^f \ar@<-.5ex>[r]_{1_X} & X \ar[r]^i & C
  } $
  is the coequalizer of $\parpairi{X}{X}{f}{1}$.
  The equivalence $\sim$ has the property:
  $\forall x_1, x_2 \in X, (x_1, x_2) \in \sim \;\Leftrightarrow\; f(x_1) = f(x_2).$
  Thus we have $[x] = \set{x' \in X \;\big|\; f(x) = f(x')}$.

>>>>

\begin{answer}
  The definition of the coequalizer follows the scheme like in the talk.
  We define the relation $R := \set{ (1_X (x), f(x) \;\big|\; x \in X} = \set{ (x, f(x) \;\big|\; x \in X} = f$ seen as a functional relation.
  Now $\sim$ is the reflexive symmetric transitive closure of $R$.
  We define $i(x) := [x] := \set{ x' \in X | x \sim x'}$
  and the quotient $C := X / \sim \;=\; i(X)$.
  Now
  $ \xymatrix{
    X \ar@<.5ex>[r]^f \ar@<-.5ex>[r]_{1_X} & X \ar[r]^i & C
  } $
  is the coequalizer of $\parpairi{X}{X}{f}{1}$.

  We discover that $(x, y) \in \sim$ means that we can ``draw a line of forward and backward steps'' between $x$ and $y$.
  A ``forward step'' means that if we have $(x, y) \in \sim$ and $f(y)=z$ then $(x, z) \in \sim$.
  A ``backward step'' means that if we have $(x, y) \in \sim$ and $f(z)=y$ then $(x, z) \in \sim$.
  We also have $(x, x) \in \sim$ for every $x \in X$.
  More formally:
  With $\hat{R} := R \cup R^{-1}$ we have
  $\sim = \bigcup_{n\in\nat} {(\hat{R})^n}$
  where $(\hat{R})^n = \underbrace{\hat{R} \of \cdots \of \hat{R}}_{\text{$n$ times}}$ is the $n$th power relation over $\hat{R}$.
\end{answer}

### 2017-08-03 sheets/talk-7/sheet-7.tex
Merge branch 'uds-psl/master'

\begin{exercise}[Limit definition]
  \begin{enumerate}
  \item Verify that terminal objects are a valid instance of the general limit definition.
  \item Verify that equalizers are a valid instance of the general limit definition.
  \end{enumerate}
\end{exercise}
\begin{answer}
  Recall that general limits are universal cones over a diagram.
  Therefore, instances of general limits are defined by the shape of their diagram.
  In the following let $\mathscr{C}$ be any category.
  \begin{enumerate}
  \item
    We choose the empty category as the shape $\bf{I}$.
    Let $D : \bf{I} \to \mathscr{C}$ a diagram. A cone over $D$ is an object $U$ with
    maps to every object in the diagram. Since the diagram is empty, the cone is just
    the object $U$.
    $U$ is universal if for any other cone $V$ in $\mathscr{C}$, there is a unique map
    $h : V \to U$.
    Any other cone $V$ is also just an object in $\mathscr{C}$.
    Therefore $U$ is a terminal object.
  \item
    As the shape $\bf{I}$ we choose the following diagram.
    \[
      \begin{tikzcd}
        I \arrow[r,shift left, "m"] \arrow[r,shift right,swap,"n"] & J
      \end{tikzcd}
    \]
    Let $D : \bf{I} \to \mathscr{C}$ a diagram.
    A cone over $D$ is an object $U$ with maps to every object in the diagram, such that all triangles commute.
    \[
      \begin{tikzcd}
        U \arrow{d}{f_I} \arrow{dr}{f_J} & \\
        D(I) \arrow[r,shift left, "Dm"] \arrow[r,shift right,swap,"Dn"] & D(J)
      \end{tikzcd}
    \]
    Specifically, the commuting condition says that $Dm \circ f_I = f_J = Dn \circ f_I$.
    Since $f_J$ is already completely determined by $f_I$, we can ignore the arrow
    and state equivalently that the following diagram commutes.
    \[
      \begin{tikzcd}
        U \arrow{r}{f_I} & D(I) \arrow[r,shift left, "Dm"] \arrow[r,shift right,swap,"Dn"] & D(J)
      \end{tikzcd}
    \]
    The arrow $f_J$ is simply left implicit.\\
    $U$ is a universal cone, i.e. a limit, if for any other cone $V$ in $\mathscr{C}$, there is a unique map
    $h : V \to U$.
    \[
      \begin{tikzcd}
        U \arrow{r}{f_I} & D(I) \arrow[r,shift left, "Dm"] \arrow[r,shift right,swap,"Dn"] & D(J) \\
        V \arrow{ur}{g_I} \arrow[dashed]{u}{h}& &
      \end{tikzcd}
    \]
    This is exactly the definition of an equalizer.
  \end{enumerate}
\end{answer}

@@

\begin{exercise}[Cone category]
  Consider an object $L \in$ \texttt{Cone(D)} that corresponds to a limit in $\mathscr{C}$ over $D$.
  What kind of object is $L$?\\
  What about an object $L' \in$ \texttt{Cone(D)} that corresponds to a colimit in $\mathscr{C}$ over $D$?
\end{exercise}
\begin{answer}
  If $L$ corresponds to a limit, then by definition any other cone with vertex $V$ in $\mathscr{C}$ has a unique map
  from $V$ to the limit that corresponds to a map in \texttt{Cone(D)}.
  Because the cones in $\mathscr{C}$ are exactly the objects in \texttt{Cone(D)},
  $L$ has a unique arrow from every object \texttt{Cone(D)}.
  Therefore $L$ is a terminal object.\\
  By duality $L'$ is an initial object.
\end{answer}

@@

\begin{exercise}[Cones as natural transformations]
  Let $\mathscr{C}$ and $\bf{I}$ be any categories, $D : \bf{I} \to \mathscr{C}$ a diagram.
  Verify that a natural transformation $\pi : \Updelta_V \to D(\bf{I})$ is a cone over $D$ with vertex $V$.
\end{exercise}
\begin{answer}
  Recall that a natural transformation $\pi : \Updelta_V \to D(\bf{I})$ has components for every object $I \in \bf{I}$
  and for every morphism $m : I \to J \in \bf{I}$ the following square commutes.
  \[
    \begin{tikzcd}
      \Updelta_V (I) \arrow{r}{\pi_I}  \arrow{d}{\Updelta_V (m)}  & D(I) \arrow{d}{D(m)}    \\
      \Updelta_V (J) \arrow{r}{\pi_J}                & D(J)
    \end{tikzcd}
  \]
  Now we unfold the definition of $\Updelta_V$.
  \[
    \begin{tikzcd}
      \Updelta_V (I) \arrow{r}{\pi_I}  \arrow{d}{\Updelta_V (m)}  & D(I) \arrow{d}{D(m)}    \\
      \Updelta_V (J) \arrow{r}{\pi_J}                & D(J)
    \end{tikzcd}
    \ = \
    \begin{tikzcd}
      V \arrow{r}{\pi_I}  \arrow{d}{1_V}  & D(I) \arrow{d}{D(m)}    \\
      V \arrow{r}{\pi_J}                & D(J)
    \end{tikzcd}
    \ = \
    \begin{tikzcd}
      V \arrow{r}{\pi_I} \arrow{dr}{\pi_J} & D(I) \arrow{d}{D(m)}    \\
                     & D(J)
    \end{tikzcd}
  \]
  To summarize, the natural transformation $\pi$ has maps $\pi_I$ for all $I \in \bf{I}$,
  and the above triangle commutes for all maps $m : I \to J$.
  This is exactly the definition of a cone with vertex $V$.
  In fact this yields an alternative equivalent definition of cones.
\end{answer}

@@

\begin{exercise}[Limit construction]
  Recall the construction:
  \[
    \begin{tikzcd}
      E \arrow{r}{e} &
      \displaystyle \prod_{I \in \bf{I}} D(I) \arrow[r,shift left, "s"] \arrow[r,shift right,swap,"t"] &
      \displaystyle \prod_{J \overset{q}{\to} K \in \bf{I}} D(K)
    \end{tikzcd}
  \]
  where $E$ is the equalizer of $s$ and $t$.
  We have shown that $E$ is a cone for any diagram $D$. Show that $E$ is also the limit,
  i.e. a universal cone over $D$.
\end{exercise}
\begin{hint}
  \begin{itemize}
  \item You may use the fact that $E$ is a cone iff $E$ is an equalizer of $s$ and $t$.
  \item You do not need the concrete definitions of $s$ and $t$.
  \end{itemize}
\end{hint}
\begin{answer}
  To show that $E$ is a universal cone, we need to show that for any other cone $F$, there is a unique 
  morphism $h : F \to E$.
  If $F$ is a cone, then it is also an equalizer of $s$ and $t$.
  \[
    \begin{tikzcd}
      E \arrow{r}{e} &
      \displaystyle \prod_{I \in \bf{I}} D(I) \arrow[r,shift left, "s"] \arrow[r,shift right,swap,"t"] &
      \displaystyle \prod_{J \overset{q}{\to} K \in \bf{I}} D(K) \\
      F \arrow{ur}{f} & &
    \end{tikzcd}
  \]
  Since $E$ is an equalizer, and $F$ is a cone over the same diagram, we have a unique map $h : F \to E$.
  \[
    \begin{tikzcd}
      E \arrow{r}{e} &
      \displaystyle \prod_{I \in \bf{I}} D(I) \arrow[r,shift left, "s"] \arrow[r,shift right,swap,"t"] &
      \displaystyle \prod_{J \overset{q}{\to} K \in \bf{I}} D(K) \\
      F \arrow{ur}{f} \arrow[dashed]{u}{h} & &
    \end{tikzcd}
  \]

>>>>

\begin{exercise}[Limit definition]
  \begin{enumerate}
  \item Verify that terminal objects are a valid instance of the general limit definition.
  \item Verify that equalizers are a valid instance of the general limit definition.
  \end{enumerate}
\end{exercise}
\begin{answer}
  Recall that general limits are universal cones over a diagram.
  Therefore, instances of general limits are defined by the shape of their diagram.
  In the following let $\mathscr{C}$ be any category.
  \begin{enumerate}
  \item
    We choose the empty category as the shape $\bf{I}$.
    Let $D : \bf{I} \to \mathscr{C}$ a diagram. A cone over $D$ is an object $U$ with
    maps to every object in the diagram. Since the diagram is empty, the cone is just
    the object $U$.
    $U$ is universal if for any other cone $V$ in $\mathscr{C}$, there is a unique map
    $h : V \to U$.
    Any other cone $V$ is also just an object in $\mathscr{C}$.
    Therefore $U$ is a terminal object.
  \item
    As the shape $\bf{I}$ we choose the following category.
    \[
      \begin{tikzcd}
        I \arrow[r,shift left, "m"] \arrow[r,shift right,swap,"n"] & J
      \end{tikzcd}
    \]
    Let $D : \bf{I} \to \mathscr{C}$ a diagram.
    A cone over $D$ is an object $U$ with maps to every object in the diagram, such that all triangles commute.
    \[
      \begin{tikzcd}
        U \arrow{d}{f_I} \arrow{dr}{f_J} & \\
        D(I) \arrow[r,shift left, "Dm"] \arrow[r,shift right,swap,"Dn"] & D(J)
      \end{tikzcd}
    \]
    Specifically, the commuting condition says that $Dm \circ f_I = f_J = Dn \circ f_I$.
    Since $f_J$ is already completely determined by $f_I$, we can ignore the arrow
    and state equivalently that the following diagram commutes.
    \[
      \begin{tikzcd}
        U \arrow{r}{f_I} & D(I) \arrow[r,shift left, "Dm"] \arrow[r,shift right,swap,"Dn"] & D(J)
      \end{tikzcd}
    \]
    The arrow $f_J$ is simply left implicit.\\
    $U$ is a universal cone, i.e. a limit, if for any other cone $V$ in $\mathscr{C}$, there is a unique map
    $h : V \to U$.
    \[
      \begin{tikzcd}
        U \arrow{r}{f_I} & D(I) \arrow[r,shift left, "Dm"] \arrow[r,shift right,swap,"Dn"] & D(J) \\
        V \arrow{ur}{g_I} \arrow[dashed]{u}{h}& &
      \end{tikzcd}
    \]
    This is exactly the definition of an equalizer.
  \end{enumerate}
\end{answer}

@@

\begin{exercise}[Cone category]
  Consider an object $L \in$ \texttt{Cone(D)} that corresponds to a limit in $\mathscr{C}$ over $D$.
  What kind of object is $L$?\\
\end{exercise}
\begin{answer}
  If $L$ corresponds to a limit, then by definition any other cone with vertex $V$ in $\mathscr{C}$ has a unique map
  from $V$ to the limit that corresponds to a map in \texttt{Cone(D)}.
  Because the cones in $\mathscr{C}$ are exactly the objects in \texttt{Cone(D)},
  $L$ has a unique arrow from every object \texttt{Cone(D)}.
  Therefore $L$ is a terminal object.\\
  There is also a dual category $\texttt{Cone(D)}^{Op}$, the category of cocones. By duality,
  an object $L'$ corresponding to a colimit in $\mathscr{C}$ is an initial object in $\texttt{Cone(D)}^{Op}$.
\end{answer}

@@

\begin{exercise}[Cones as natural transformations]
  Let $\mathscr{C}$ and $\bf{I}$ be any categories, $D : \bf{I} \to \mathscr{C}$ a diagram.
  Verify that a natural transformation $\pi : \Updelta_V \to D$ is a cone over $D$ with vertex $V$.
\end{exercise}
\begin{answer}
  Recall that a natural transformation $\pi : \Updelta_V \to D$ has components for every object $I \in \bf{I}$
  and for every morphism $m : I \to J \in \bf{I}$ the following square commutes.
  \[
    \begin{tikzcd}
      \Updelta_V (I) \arrow{r}{\pi_I}  \arrow{d}{\Updelta_V (m)}  & D(I) \arrow{d}{D(m)}    \\
      \Updelta_V (J) \arrow{r}{\pi_J}                & D(J)
    \end{tikzcd}
  \]
  Now we unfold the definition of $\Updelta_V$.
  \[
    \begin{tikzcd}
      \Updelta_V (I) \arrow{r}{\pi_I}  \arrow{d}{\Updelta_V (m)}  & D(I) \arrow{d}{D(m)}    \\
      \Updelta_V (J) \arrow{r}{\pi_J}                & D(J)
    \end{tikzcd}
    \ = \
    \begin{tikzcd}
      V \arrow{r}{\pi_I}  \arrow{d}{1_V}  & D(I) \arrow{d}{D(m)}    \\
      V \arrow{r}{\pi_J}                & D(J)
    \end{tikzcd}
    \ = \
    \begin{tikzcd}
      V \arrow{r}{\pi_I} \arrow{dr}{\pi_J} & D(I) \arrow{d}{D(m)}    \\
                     & D(J)
    \end{tikzcd}
  \]
  To summarize, the natural transformation $\pi$ has maps $\pi_I$ for all $I \in \bf{I}$,
  and the above triangle commutes for all maps $m : I \to J$.
  This is exactly the definition of a cone with vertex $V$.
  In fact this yields an alternative equivalent characterization of cones.
\end{answer}

@@

\begin{exercise}[Limit construction]
  Recall the construction:
  \[
    \begin{tikzcd}
      E \arrow{r}{e} &
      \displaystyle \prod_{I \in \bf{I}} D(I) \arrow[r,shift left, "s"] \arrow[r,shift right,swap,"t"] &
      \displaystyle \prod_{J \overset{q}{\to} K \in \bf{I}} D(K)
    \end{tikzcd}
  \]
  where $E$ is the equalizer of $s$ and $t$.
  We have shown that $E$ is a cone for any diagram $D$. Show that $E$ is also the limit,
  i.e. a universal cone over $D$.
\end{exercise}
\begin{hint}
  \begin{itemize}
  \item You may use the fact that $E$ is a cone iff $E$ is a fork of $s$ and $t$.
  \item You do not need the concrete definitions of $s$ and $t$.
  \end{itemize}
\end{hint}
\begin{answer}
  To show that $E$ is a universal cone, we need to show that for any other cone $F$, there is a unique 
  morphism $h : F \to E$.
  If $F$ is a cone, then it is also a fork of $s$ and $t$.
  \[
    \begin{tikzcd}
      E \arrow{r}{e} &
      \displaystyle \prod_{I \in \bf{I}} D(I) \arrow[r,shift left, "s"] \arrow[r,shift right,swap,"t"] &
      \displaystyle \prod_{J \overset{q}{\to} K \in \bf{I}} D(K) \\
      F \arrow{ur}{f} & &
    \end{tikzcd}
  \]
  Since $E$ is an equalizer, and $F$ is a cone over the same diagram, we have a unique map $h : F \to E$.
  \[
    \begin{tikzcd}
      E \arrow{r}{e} &
      \displaystyle \prod_{I \in \bf{I}} D(I) \arrow[r,shift left, "s"] \arrow[r,shift right,swap,"t"] &
      \displaystyle \prod_{J \overset{q}{\to} K \in \bf{I}} D(K) \\
      F \arrow{ur}{f} \arrow[dashed]{u}{h} & &
    \end{tikzcd}
  \]
